# Lecture 02

## What proportion of the surface is covered with water? 

## Bayesian data analysis

For each possible explanation of the data



## Globe tossing

### 


* No point estimate

The distribution is the estimate

Always use the entire distribtution 

* No one true interval 

Intervals communicate shape of posterior. 

50% and 89% are just arbitrary numbers used. 

95% is obvious superstition. 
Nothing magical happens at the boundary. 

None of this is logical. It's just conventions. 


### Letters from my reviewers

"The authors uses these cute *89% intervals*, but we need to see the *95% intervals*
so we can tell whether any of the effects are *robust*. "

> That an arbitrary interval contains an arbitrary value is not meaningful. Use the whole distribtution. The intervals are just for a short summary of the distribution. 


## Formalities


The observations (data) and explanations (parametrers) are variables. 

For each variable, must say how it is generated. 

1. Data: W and L, the number of water and land obervations

$$
\operatorname{Pr}(W, L \mid p)=\frac{(W+L) !}{W ! L !} p^{W}(1-p)^{L}
$$
$$
\operatorname{Pr} (W, L \mid p) = \frac{(W + L)!} {W! L!} p^{W} (1-p)^{L}
$$


Binomial probability function:

$$
\text { dbinom (} W, W+L, p)
$$

```{r}
# x: vectior of quarantines
# size: number of observations
# p vector of probabilities
# p(x) = choose(n, x) p^x (1-p)^(n-x)
# for x = 0, …, n. Note that binomial coefficients can be computed by choose in R.
dbinom(6, 9, 0.7)
```

2. Parameters: p, the proportion of water on the globe

Relative plausibility of each possible $p$

$$
\operatorname{Pr} (p) = \frac{1} {1-0} = 1
$$


Posterior is (normalized) product

$$
\operatorname{Pr} (p \mid W, L) = \frac{\operatorname{Pr} (W, L \mid p)\operatorname{Pr}(p)} {\operatorname{Pr} (W, L)} 
$$

For each possible value of $p$, to get the probability of that $p$ conditional on W and L 

$\operatorname{Pr} (W, L)$ is the normalization constant

3. With number

For each possible value of p, compute 



## Grid approximation


```{r}
p_grid <- seq(from = 0, to = 1, length.out = 1000)
prob_p <- rep(1, 1000)
prob_data <- dbinom(6, 9, prob = p_grid)
posterior <- prob_data * prob_p
posterior_normalized <- posterior / sum(posterior)
```

```{r}
par(mfrow=c(2,2))
plot(p_grid)
plot(prob_p)
plot(prob_data)
## Probability distribution
plot(posterior_normalized)
```

```{r}
p_grid <- seq(from = 0, to = 1, length.out = 1000)
prob_p <- dbeta(p_grid, 3, 1)
prob_data <- dbinom(6, 9, prob = p_grid)
posterior <- prob_data * prob_p
posterior_normalized_beta <- posterior / sum(posterior)
par(mfrow=c(2,2))
plot(p_grid)
plot(prob_p)
plot(prob_data)
## Probability distribution
plot(posterior_normalized, col="gray", 
     ylim=c(0,max(posterior_normalized, posterior_normalized_beta)))
lines(posterior_normalized_beta)
```


## Reference

https://www.youtube.com/watch?v=guTdrfycW2Q&list=PLDcUM9US4XdMROZ57-OIRtIK0aOynbgZN&index=2
