---
title: "A Book"
author: "Frida Gomam"
site: bookdown::bookdown_site
documentclass: book
output:
  bookdown::gitbook: default
  #bookdown::pdf_book: default
---

# Hello World

Hi.

Bye.

<!-- If you need PDF output, uncomment bookdown::pdf_book above in YAML. You will need a LaTeX installation, e.g., https://yihui.name/tinytex/ -->

<!--chapter:end:index.Rmd-->

# Lecture 07



## Copernican Model

```{r}

library(rethinking)
library(animation)
```



## Reference 

https://www.youtube.com/watch?v=odGAAJDlgp8


<!--chapter:end:07_lecture07.Rmd-->

# Lecture 09




## Generative model 


```{r}
library(simcausal)
N <- 1000

G <- sample(1:2, size = N, replace = TRUE)

D <- rbern(N, ifelse(G==1, 0.3, 0.8)) + 1 

accept_rate <- matrix(c(0.05, 0.2, 0.1, 0.3), nrow = 2)

A <- rbern(N, accept_rate[D, G])
```

```{r}
table(G, D)
```

```{r}
table(G, A)
```


## 

Is a start, but lots missing

Addmission rate usaully depends upon size of applicant pool, distribution of qualifications


In principle, should sample applicant pool and then sort to select admissions

Rates are conditional on structure of applicant pool. 



## Generalized Linear Models


Linear Models



Generalized linear modesl: 


Expoected value is some fuction of an additive combination of parameters




## Links and inverse links






## Reference

https://www.youtube.com/watch?v=nPi5yGbfxuo


<!--chapter:end:09_lecture9.Rmd-->

